---
title: "SSalgado_tidy_a"
author: "Samantha Salgado"
date: "5/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```

#Q1
```{r}
art <- read_csv("https://github.com/MuseumofModernArt/collection/blob/master/Artworks.csv?raw=true")
str(art)
```
#Q2
```{r}
art <- separate(art, ConstituentID, into = c("firstID", "secondID", sep = ",",
                                             convert = TRUE))
select(art, firstID, secondID)
```

#Q3
```{r}
select(art, firstID, secondID, ObjectID, Title) %>% slice(110:115)
```

#part 2
#one observation per objectID: This one checks out and is good!
```{r}
nrow(art) - n_distinct(art$ObjectID)
```

#This filters out any duplicates
```{r}
art2 <- filter(art, firstID == secondID) %>% nrow

```

#this gives an NA to any duplicate artists. 
```{r}
art1 <- mutate(art, secondID = ifelse(firstID==secondID, NA, secondID))
```
#the data set is also has large names for the museums. Maybe acronyms could help? 
#would the data tables really need all the information it provides? maybe can be split up into smaller data tables that can call on each other.


#the above gives a data frame with a comma as a column, and below will remove the NA columns of secondID and the comma column
```{r}
art <-filter(art, firstID != NA)
head(art)

drop <- c("secondID", ",")

```

